---
title: "R basics"
author: "Yuanhang Liu"
date: "Monday, Nov 30, 2015"
output: 
  BiocStyle::html_document:
    toc: true
---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

# Installation

Please install several required packages if you haven't install them already:

```{r, eval=FALSE}
install.packages('ggplot2')
install.packages('lattice')
```

```{r, message=FALSE}
library(lattice)
library(ggplot2)
```


# Basic operations in R

You can do use R as one calculator

```{r}
1 + 2*3 + 4
```

You can also generate one array:

```{r}
c(0, 1, 1, 2, 3, 5, 8)
```

Or use sequence operator:

```{r}
1:20
```

You can also enter expression of strings:

```{r}
"Hello world."
```

# Functions

Functions are basically one or multiple operations 

```{r, eval=FALSE}
f(argument1, argument2, ...)
```

This function compute log2 of 64:

```{r}
log(x=64, base=2)
```

A binary operator is also a function:

```{r}
5 + 4
```

# Data structures

In R, you can construct more complicated data structures than just vectors. You can define one array, a multi-dimensional vector, like this:

```{r}
a <- array(c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12), dim=c(3, 4))
a
a[2,2]
a[2,]
a[1:2,]
a[c(1,3),c(2,4)]
```

One array can be more than two dimensions:

```{r}
w <- array(c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18),dim=c(3,3,2))
w
w[1,2,2]
```

A matrix is basically a two dimensional array:

```{r}
m <- matrix(data=c(1,2,3,4,5,6,7,8,9,10,11,12),nrow=3,ncol=4)
m
m[2,3]
```

In all the examples above, we’ve just looked at data structures based on a single underlying data type. If you want to construct more complicated structures with multiple data types, you will need to construct list or data.frame:

A list can be constructed by:

```{r}
e <- list(thing="hat", size="8.25")
e
e$thing
e[1]
e[[1]]
```

A data.frame is basically a list that contains multiple named vectors that are the same lengh:

```{r}
teams <- c("PHI","NYM","FLA","ATL","WSN")
w <- c(92, 89, 94, 72, 59)
l <- c(70, 73, 77, 90, 102)
nleast <- data.frame(teams,w,l)
nleast
```

# Data structures in R

R is an object-oriented language. Every object in R has a type. Additionally, every object in R is a member of a class. We have already encountered several different classes: character vectors, numeric vectors, data frames, lists, and arrays.

```{r}
class(TRUE)
class('TRUE')
class(4)
class(cars)
```

```{r}
17 + 6
as.Date("2009-09-08") + 7
```

# Models and formulas

R has a special notation for describing relationships between variables. 

```{r}
str(cars)
cars.lm <- lm(formula=dist~speed,data=cars)
cars.lm
summary(cars.lm)
```

# Charts and graphics

R mainly has three different styles of plotting system: base plot, lattice and ggplot2:

A base plot can be initiated by:

```{r}
plot(dist~speed, type = 'p', data = cars, main = 'base plot test')
```

For lattice

```{r}
xyplot(dist~speed, data=cars, main = 'lattice test')
```

For ggplot2:

```{r}
ggplot() +
  geom_point(aes(x = speed, y = dist), data = cars) +
  ggtitle('ggplot2 test')
```

# Getting help

```{r, eval=FALSE}
?ggplot
```